---
title: "Assignment 9: Graduate School Admissions"
author: "Mason Hayes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(broom)
```

## Introduction

This report explores predictors of graduate school admission using GRE score, GPA, and undergraduate institution rank. Because the response variable, admission, is binary, logistic regression is used to model the probability of admission.

# Load Data
```{r}
admissions <- read_csv("GradSchool_Admissions.csv")
glimpse(admissions)
summary(admissions)
```

# Data Cleaning
```{r}
admissions <- admissions %>%
  mutate(rank = factor(rank))

glimpse(admissions)
```
# Exploratory Data Analysis

## Plot 1: Admission Counts
```{r}
ggplot(admissions, aes(x = factor(admit))) +
  geom_bar() +
  labs(
    title = "Admission Outcomes",
    x = "Admit (0 = No, 1 = Yes)",
    y = "Count"
  )
```

## Plot 2: GPA vs admission
```{r}
ggplot(admissions, aes(x = factor(admit), y = gpa)) +
  geom_boxplot() +
  labs(
    title = "GPA by Admission Outcome",
    x = "Admit",
    y = "GPA"
  )
```

## Plot 3: GRE vs admission
```{r}
ggplot(admissions, aes(x = factor(admit), y = gre)) +
  geom_boxplot() +
  labs(
    title = "GRE by Admission Outcome",
    x = "Admit",
    y = "GRE"
  )
```

## Plot 4: Rank vs admission
```{r}
ggplot(admissions, aes(x = rank, fill = factor(admit))) +
  geom_bar(position = "fill") +
  labs(
    title = "Admission Rate by Rank",
    x = "Rank",
    y = "Proportion",
    fill = "Admit"
  )
```

## Interpretation
The exploratory analysis suggests that GPA is higher on average for admitted students compared to non-admitted students. GRE scores also appear slightly higher among admitted students, though the difference is less pronounced. Admission rates vary by undergraduate institution rank, with higher-ranked institutions showing higher proportions of admitted students.

## Logistic Regression Model
```{r}
model <- glm(admit ~ gre + gpa + rank, 
             data = admissions, 
             family = "binomial")

summary(model)
```

## Model Interpretation
The logistic regression model shows that GPA is a strong positive predictor of admission, meaning higher GPA increases the likelihood of being admitted. GRE scores also have a positive effect, though smaller. Undergraduate institution rank has a negative relationship with admission, indicating that students from lower-ranked institutions are less likely to be admitted.

## Predicted Probabilities
```{r}
admissions$prob <- predict(model, type = "response")
head(admissions)
```